# CAD optimal operating point II {#cad-optim-op-point-2}

---
output:
  rmarkdown::pdf_document:
    fig_caption: yes        
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval.after = "fig.cap"
)

library(kableExtra)
library(gridExtra)
library(dplyr)
```



```{r, include=FALSE}
source("R/collapse_rows_df.R", local = knitr::knit_global())
```



```{r, echo=FALSE}
source(here::here("R/optim-op-point/optim-op-point-helpers.R"))
```



## Varying $\nu$ optimizations{#optim-op-point-vary-nu}

For $\mu = 2$ and $\lambda = 1$ optimizations were performed for $\nu = 0.6, 0.7, 0.8, 0.9$. 



```{r, echo = TRUE}
muArr <- c(2)
lambdaArr <- c(1)
nuArr <- c(0.6, 0.7, 0.8, 0.9)
lesDistr <- c(0.5, 0.5)
relWeights <- c(0.5, 0.5)
```


```{r, include=FALSE}
source("R/optim-op-point/doOneTable.R", local = knitr::knit_global())
```



### Summary table


```{r optim-op-point-table-vary-nu-contents, echo=FALSE}
x12 <- doOneTable(nuArr, zetaOptArr, wAfrocArr, rocArr, nlfOptArr, llfOptArr, 1,1,1:4, labelIndx = 3)
```




```{r optim-op-point-table-vary-nu, echo=FALSE}
df <- collapse_rows_df(x12,FOM)

captionStr <- "Summary of optimization results for $\\mu = 2$, $\\lambda = 1$ and varying $\\nu$."
  
kbl(df, caption = captionStr, booktabs = TRUE, escape = FALSE) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10) 
```




### FROC



```{r echo=FALSE}
plotArr = plotFroc(muArr, lambdaArr, nuArr)
```




```{r optim-op-point-vary-nu-froc, fig.cap = "Varying $\\nu$ FROC plots with superimposed operating points. The red dot corresponds to wAFROC-AUC optimization and the black dot to Youden-index optimization. The values of $\\nu$ are: top-left $\\nu = 0.6$, top-right $\\nu = 0.7$, bottom-left $\\nu = 0.8$ and bottom-right $\\nu = 0.9$. Each red dot is above the corresponding black dot and their separation increases as $\\nu$ increases, i.e., as CAD performance increases.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```



### wAFROC


 
```{r echo=FALSE}
plotArr = plotwAfroc(muArr, lambdaArr, nuArr, lesDistr, relWeights)
```




```{r optim-op-point-vary-nu-wafroc, fig.cap = "Varying $\\nu$ wAFROC plots for the two optimization methods. The color coding is as in previous figures. The values of $\\nu$ are: top-left $\\nu = 0.6$, top-right $\\nu = 0.7$, bottom-left $\\nu = 0.8$ and bottom-right $\\nu = 0.9$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```



### ROC



```{r echo=FALSE}
plotArr = plotRoc(muArr, lambdaArr, nuArr, lesDistr, relWeights)
```







```{r optim-op-point-vary-nu-roc, fig.cap = "Varying $\\nu$ ROC plots for the two optimization methods. The color coding is as in previous figures. The values of $\\nu$ are: top-left $\\nu = 0.6$, top-right $\\nu = 0.7$, bottom-left $\\nu = 0.8$ and bottom-right $\\nu = 0.9$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```





## Varying $\mu$ optimizations{#optim-op-point-vary-mu}

For $\lambda = 1$ and $\nu = 0.9$ optimizations were performed for $\mu = 1, 2, 3, 4$. 



```{r}
muArr <- c(1, 2, 3, 4)
lambdaArr <- 1
nuArr <- 0.9
```



```{r, include=FALSE}
source("R/optim-op-point/doOneTable.R", local = knitr::knit_global())
```



### Summary table



```{r optim-op-point-table-vary-mu-contents, echo=FALSE}
x12 <- doOneTable(muArr, zetaOptArr, wAfrocArr, rocArr, nlfOptArr, llfOptArr, 1:4,1,1, labelIndx = 1)
```





```{r optim-op-point-table-vary-mu, echo=FALSE}
df <- collapse_rows_df(x12,FOM)

captionStr <- "Summary of optimization results for $\\lambda = 1$, $\\nu = 0.9$ and varying $\\mu$."
  
kbl(df, caption = captionStr, booktabs = TRUE, escape = FALSE) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10) 
```



### FROC


```{r echo=FALSE}
plotArr = plotFroc(muArr, lambdaArr, nuArr)
```




```{r optim-op-point-vary-mu-froc, fig.cap = "Varying $\\mu$ FROC plots with superimposed operating points. The red dot corresponds to wAFROC-AUC optimization and the black dot to Youden-index optimization. The values of $\\mu$ are: top-left $\\mu = 1$, top-right $\\mu = 2$, bottom-left $\\mu = 3$ and bottom-right $\\mu = 4$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```



### wAFROC


```{r echo=FALSE}
plotArr = plotwAfroc(muArr, lambdaArr, nuArr, lesDistr, relWeights)
```




```{r optim-op-point-vary-mu-wafroc, fig.cap = "Varying $\\mu$ wAFROC plots for the two optimization methods. The color coding is as in previous figures. The values of $\\mu$ are: top-left $\\mu = 1$, top-right $\\mu = 2$, bottom-left $\\mu = 3$ and bottom-right $\\mu = 4$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```




### ROC


```{r echo=FALSE}
plotArr = plotRoc(muArr, lambdaArr, nuArr, lesDistr, relWeights)
```





```{r optim-op-point-vary-mu-roc, fig.cap = "Varying $\\mu$ ROC plots for the two optimization methods. The color coding is as in previous figures. The values of $\\mu$ are: top-left $\\mu = 1$, top-right $\\mu = 2$, bottom-left $\\mu = 3$ and bottom-right $\\mu = 4$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```




## Limiting cases {#optim-op-point-limiting-situations}

### High performance varying $\mu$ {#optim-op-point-high-performance-vary-mu}


```{r}
muArr <- c(2, 3, 4, 5)
nuArr <- c(0.9)
lambdaArr <- c(1)
```



```{r, include=FALSE}
source("R/optim-op-point/doOneTable.R", local = knitr::knit_global())
```



#### Summary table



```{r optim-op-point-high-performance-vary-mu-table-contents, echo=FALSE}
x12 <- doOneTable(muArr, zetaOptArr, wAfrocArr, rocArr, nlfOptArr, llfOptArr, 1:4,1,1, labelIndx = 1)
```



```{r optim-op-point-high-performance-vary-mu-table, echo=FALSE}
df <- collapse_rows_df(x12,FOM)

captionStr <- "High performance summary of optimization results for $\\lambda = 1$ and $\\nu = 0.9$ and varying $\\mu$."
  
kbl(df, caption = captionStr, booktabs = TRUE, escape = FALSE) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10) 
```



#### FROC



```{r echo=FALSE}
plotArr = plotFroc(muArr, lambdaArr, nuArr)
```




```{r optim-op-point-high-performance-vary-mu-froc, fig.cap = "High performance varying $\\mu$ FROC plots with superimposed operating points. The red dot corresponds to wAFROC-AUC optimization and the black dot to Youden-index optimization. The values of $\\mu$ are: top-left $\\mu = 2$, top-right $\\mu = 3$, bottom-left $\\mu = 4$ and bottom-right $\\mu = 5$.", fig.show='hold', echo=FALSE}
# DpcBugFix 8/28/22
# in next line the generating code has been altered to suppress title on each plot; TBA; TODOLAST 
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```




#### wAFROC



```{r echo=FALSE}
plotArr = plotwAfroc(muArr, lambdaArr, nuArr, lesDistr, relWeights)
```



```{r optim-op-point-high-performance-vary-mu-wafroc, fig.cap = "High performance varying $\\mu$ wAFROC plots for the two optimization methods. The color coding is as in previous figures. The values of $\\mu$ are: top-left $\\mu = 2$, top-right $\\mu = 3$, bottom-left $\\mu = 4$ and bottom-right $\\mu = 5$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
#grid.arrange(plotArr[[5]], plotArr[[6]], plotArr[[7]], plotArr[[8]], nrow = 2, ncol = 2)
```




#### ROC



```{r echo=FALSE}
plotArr = plotRoc(muArr, lambdaArr, nuArr, lesDistr, relWeights)
```




```{r optim-op-point-high-performance-vary-mu-roc, fig.cap = "High performance varying $\\mu$ ROC plots for the two optimization methods. The color coding is as in previous figures. The values of $\\mu$ are: top-left $\\mu = 2$, top-right $\\mu = 3$, bottom-left $\\mu = 4$ and bottom-right $\\mu = 5$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```


### Low performance varying $\mu$ {#optim-op-point-low-performance-vary-mu}


```{r}
muArr <- c(1, 2, 3, 4)
nuArr <- c(0.1)
lambdaArr <- c(10)
```




```{r, include=FALSE}
source("R/optim-op-point/doOneTable.R", local = knitr::knit_global())
```



#### Summary table


```{r optim-op-point-low-performance-vary-mu-table-contents, echo=FALSE}
x12 <- doOneTable(muArr, zetaOptArr, wAfrocArr, rocArr, nlfOptArr, llfOptArr, 1:4,1,1, labelIndx = 1)
```



```{r optim-op-point-low-performance-vary-mu-table, echo=FALSE}
df <- collapse_rows_df(x12,FOM)

captionStr <- "Low performance summary of optimization results for $\\lambda = 10$ and $nu = 0.1$ and varying $\\mu$. Column labeling as in previous tables."
  
kbl(df, caption = captionStr, booktabs = TRUE, escape = FALSE) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10) 
```


#### FROC



```{r echo=FALSE}
plotArr = plotFroc(muArr, lambdaArr, nuArr)
```




```{r optim-op-point-low-performance-vary-mu, fig.cap = "Low performance varying $\\mu$ FROC plots with superimposed operating points. The red dot corresponds to wAFROC-AUC optimization and the black dot to Youden-index optimization. The values of $\\mu$ are: top-left $\\mu = 1$, top-right $\\mu = 2$, bottom-left $\\mu = 3$ and bottom-right $\\mu = 4$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```




#### wAFROC



```{r echo=FALSE}
plotArr = plotwAfroc(muArr, lambdaArr, nuArr, lesDistr, relWeights)
```



```{r optim-op-point-low-performance-vary-mu-wafroc, fig.cap = "Low performance varying $\\mu$ wAFROC plots for the two optimization methods. The color coding is as in previous figures. The values of $\\mu$ are: top-left $\\mu = 1$, top-right $\\mu = 2$, bottom-left $\\mu = 3$ and bottom-right $\\mu = 4$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
#grid.arrange(plotArr[[5]], plotArr[[6]], plotArr[[7]], plotArr[[8]], nrow = 2, ncol = 2)
```



#### ROC



```{r echo=FALSE}
plotArr = plotRoc(muArr, lambdaArr, nuArr, lesDistr, relWeights)
```




```{r optim-op-point-low-performance-vary-mu-roc, fig.cap = "Low performance varying $\\mu$ ROC plots for the two optimization methods. The color coding is as in previous figures. The values of $\\mu$ are: top-left $\\mu = 1$, top-right $\\mu = 2$, bottom-left $\\mu = 3$ and bottom-right $\\mu = 4$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
#grid.arrange(plotArr[[5]], plotArr[[6]], plotArr[[7]], plotArr[[8]], nrow = 2, ncol = 2)
```




### High performance varying $\lambda$ {#optim-op-point-high-performance-vary-lambda}


```{r}
muArr <- c(4)
nuArr <- c(0.9)
lambdaArr <- c(1,2,5,10)
```



```{r, include=FALSE}
source("R/optim-op-point/doOneTable.R", local = knitr::knit_global())
```





#### Summary table


```{r optim-op-point-high-performance-vary-lambda-table-contents, echo=FALSE}
x12 <- doOneTable(lambdaArr, zetaOptArr, wAfrocArr, rocArr, nlfOptArr, llfOptArr, 1,1:4,1, labelIndx = 2)
```



```{r optim-op-point-high-performance-vary-lambda-table, echo=FALSE}
df <- collapse_rows_df(x12,FOM)

captionStr <- "Summary of optimization results for $\\mu = 4$, $nu = 0.9$ and varying $\\lambda$. Column labeling as in previous tables."
  
kbl(df, caption = captionStr, booktabs = TRUE, escape = FALSE) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10) 
```


#### FROC




```{r echo=FALSE}
plotArr = plotFroc(muArr, lambdaArr, nuArr)
```



```{r optim-op-point-high-performance-vary-lambda-froc, fig.cap = "High performance varying $\\lambda$ FROC plots with superimposed operating points. The red dot corresponds to wAFROC-AUC optimization and the black dot to Youden-index optimization. The values of $\\lambda$ are: top-left $\\lambda = 1$, top-right $\\lambda = 2$, bottom-left $\\lambda = 5$ and bottom-right $\\lambda = 10$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
#grid.arrange(plotArr[[5]], plotArr[[6]], plotArr[[7]], plotArr[[8]], nrow = 2, ncol = 2)
```



#### wAFROC



```{r echo=FALSE}
plotArr = plotwAfroc(muArr, lambdaArr, nuArr, lesDistr, relWeights)
```





```{r optim-op-point-high-performance-vary-lambda-wafroc, fig.cap = "High performance varying $\\lambda$ wAFROC plots for the two optimization methods. The color coding is as in previous figures. The values of $\\lambda$ are: top-left $\\lambda = 1$, top-right $\\lambda = 2$, bottom-left $\\lambda = 5$ and bottom-right $\\lambda = 10$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```



#### ROC



```{r echo=FALSE}
plotArr = plotRoc(muArr, lambdaArr, nuArr, lesDistr, relWeights)
```




```{r optim-op-point-high-performance-vary-lambda-roc, fig.cap = "High performance varying $\\lambda$ ROC plots for the two optimization methods. The color coding is as in previous figures. The values of $\\lambda$ are: top-left $\\lambda = 1$, top-right $\\lambda = 2$, bottom-left $\\lambda = 5$ and bottom-right $\\lambda = 10$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```


### Low performance varying $\lambda$ {#optim-op-point-low-performance-vary-lambda}


```{r}
muArr <- c(1)
nuArr <- c(0.2)
lambdaArr <- c(1, 2, 5, 10)
```



```{r, include=FALSE}
source("R/optim-op-point/doOneTable.R", local = knitr::knit_global())
```




#### Summary table



```{r optim-op-point-low-performance-vary-lambda-table-contents, echo=FALSE}
x12 <- doOneTable(lambdaArr, zetaOptArr, wAfrocArr, rocArr, nlfOptArr, llfOptArr, 1,1:4,1, labelIndx = 2)
```



```{r optim-op-point-low-performance-vary-lambda-table, echo=FALSE}
df <- collapse_rows_df(x12,FOM)

captionStr <- "Summary of optimization results for $\\mu = 1$, $\\nu = 0.2$ and varying $\\lambda$. Column labeling as in previous tables."
  
kbl(df, caption = captionStr, booktabs = TRUE, escape = FALSE) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10) 
```



#### FROC



```{r echo=FALSE}
plotArr = plotFroc(muArr, lambdaArr, nuArr)
```




```{r optim-op-point-low-performance-vary-lambda-froc, fig.cap = "Low performance varying $\\lambda$ FROC plots with superimposed operating points. The red dot corresponds to wAFROC-AUC optimization and the black dot to Youden-index optimization. The values of $\\lambda$ are: top-left $\\lambda = 1$, top-right $\\lambda = 2$, bottom-left $\\lambda = 5$ and bottom-right $\\lambda = 10$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```



#### wAFROC



```{r echo=FALSE}
plotArr = plotwAfroc(muArr, lambdaArr, nuArr, lesDistr, relWeights)
```







```{r optim-op-point-low-performance-vary-lambda-wafroc, fig.cap = "Low performance varying $\\lambda$ wAFROC plots for the two optimization methods. The color coding is as in previous figures. The values of $\\lambda$ are: top-left $\\lambda = 1$, top-right $\\lambda = 2$, bottom-left $\\lambda = 5$ and bottom-right $\\lambda = 10$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```



#### ROC



```{r echo=FALSE}
plotArr = plotRoc(muArr, lambdaArr, nuArr, lesDistr, relWeights)
```




```{r optim-op-point-low-performance-vary-lambda-roc, fig.cap = "Low performance varying $\\lambda$ ROC plots for the two optimization methods. The color coding is as in previous figures. The values of $\\lambda$ are: top-left $\\lambda = 1$, top-right $\\lambda = 2$, bottom-left $\\lambda = 5$ and bottom-right $\\lambda = 10$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```



### High performance varying $\nu$ {#optim-op-point-high-performance-vary-nu}


```{r}
muArr <- c(4)
lambdaArr <- c(1)
nuArr <- c(0.6, 0.7, 0.8, 0.9)
```



```{r, include=FALSE}
source("R/optim-op-point/doOneTable.R", local = knitr::knit_global())
```




#### Summary table



```{r optim-op-point-high-performance-vary-nu-table-contents, echo=FALSE}
x12 <- doOneTable(nuArr, zetaOptArr, wAfrocArr, rocArr, nlfOptArr, llfOptArr, 1,1,1:4, labelIndx = 3)
```



```{r optim-op-point-high-performance-vary-nu-table, echo=FALSE}
df <- collapse_rows_df(x12,FOM)

captionStr <- "Summary of optimization results for $\\mu = 4$, $\\lambda = 1$ and varying $\\nu$. Column labeling as in previous tables."
  
kbl(df, caption = captionStr, booktabs = TRUE, escape = FALSE) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10) 
```



#### FROC



```{r echo=FALSE}
plotArr = plotFroc(muArr, lambdaArr, nuArr)
```




```{r optim-op-point-high-performance-vary-nu-froc, fig.cap = "High performance varying $\\nu$ FROC plots with superimposed operating points. The red dot corresponds to wAFROC-AUC optimization and the black dot to Youden-index optimization. The values of $\\nu$ are: top-left $\\nu = 0.6$, top-right $\\nu = 0.7$, bottom-left $\\nu = 0.8$ and bottom-right $\\nu = 0.9$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```



#### wAFROC



```{r echo=FALSE}
plotArr = plotwAfroc(muArr, lambdaArr, nuArr, lesDistr, relWeights)
```






```{r optim-op-point-high-performance-vary-nu-wafroc, fig.cap = "High performance varying $\\nu$ wAFROC plots for the two optimization methods. The color coding is as in previous figures. The values of $\\nu$ are: top-left $\\nu = 0.6$, top-right $\\nu = 0.7$, bottom-left $\\nu = 0.8$ and bottom-right $\\nu = 0.9$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```



#### ROC



```{r echo=FALSE}
plotArr = plotRoc(muArr, lambdaArr, nuArr, lesDistr, relWeights)
```




```{r optim-op-point-high-performance-vary-nu-roc, fig.cap = "High performance varying $\\nu$ ROC plots for the two optimization methods. The color coding is as in previous figures. The values of $\\nu$ are: top-left $\\nu = 0.6$, top-right $\\nu = 0.7$, bottom-left $\\nu = 0.8$ and bottom-right $\\nu = 0.9$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```


### Low performance varying $\nu$ {#optim-op-point-low-performance-vary-nu}


```{r}
muArr <- c(1)
lambdaArr <- c(10)
nuArr <- c(0.1, 0.2, 0.3, 0.4)
```



```{r, include=FALSE}
source("R/optim-op-point/doOneTable.R", local = knitr::knit_global())
```



#### Summary table



```{r optim-op-point-low-performance-vary-nu-table-contents, echo=FALSE}
x12 <- doOneTable(nuArr, zetaOptArr, wAfrocArr, rocArr, nlfOptArr, llfOptArr, 1,1,1:4, labelIndx = 3)
```



```{r optim-op-point-low-performance-vary-nu-table, echo=FALSE}
df <- collapse_rows_df(x12,FOM)

captionStr <- "Summary of optimization results for $\\mu = 1$, $\\lambda = 10$ and varying $\\nu$. Column labeling as in previous tables."
  
kbl(df, caption = captionStr, booktabs = TRUE, escape = FALSE) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10) 
```



#### FROC



```{r echo=FALSE}
plotArr = plotFroc(muArr, lambdaArr, nuArr)
```




```{r optim-op-point-low-performance-vary-nu-froc, fig.cap = "Low performance varying $\\nu$ FROC plots with superimposed operating points. The red dot corresponds to wAFROC-AUC optimization and the black dot to Youden-index optimization. The values of $\\nu$ are: top-left $\\nu = 0.1$, top-right $\\nu = 0.2$, bottom-left $\\nu = 0.3$ and bottom-right $\\nu = 0.4$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```



#### wAFROC



```{r echo=FALSE}
plotArr = plotwAfroc(muArr, lambdaArr, nuArr, lesDistr, relWeights)
```







```{r optim-op-point-low-performance-vary-nu-wafroc, fig.cap = "Low performance varying $\\nu$ wAFROC plots for the two optimization methods. The color coding is as in previous figures. The values of $\\nu$ are: top-left $\\nu = 0.1$, top-right $\\nu = 0.2$, bottom-left $\\nu = 0.3$ and bottom-right $\\nu = 0.4$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```



#### ROC



```{r echo=FALSE}
plotArr = plotRoc(muArr, lambdaArr, nuArr, lesDistr, relWeights)
```




```{r optim-op-point-low-performance-vary-nu-roc, fig.cap = "Low performance varying $\\nu$ ROC plots for the two optimization methods. The color coding is as in previous figures. The values of $\\nu$ are: top-left $\\nu = 0.1$, top-right $\\nu = 0.2$, bottom-left $\\nu = 0.3$ and bottom-right $\\nu = 0.4$.", fig.show='hold', echo=FALSE}
grid.arrange(plotArr[[1]], plotArr[[2]], plotArr[[3]], plotArr[[4]], nrow = 2, ncol = 2)
```



