# How to analyze a dataset with zero non-diseased cases  {#analyze-diseased-only-dataset}

---
output:
  rmarkdown::pdf_document:
    fig_caption: yes        
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval.after = "fig.cap"
)

library(kableExtra)
library(gridExtra)
library(abind)
library(dplyr)
```




## TBA How much finished {#analyze-diseased-only-dataset-how-much-finished}
0%


## The problem {#analyze-diseased-only-dataset-methods}

How to analyze $K_1 = 0$ datasets.


## TBA Handling diseased-only datasets {#diseased-only-dataset}

ROC-like plot of TPF vs. FPF1 is possible. Can create a ROC-like dataset with equal number of "non-diseased" and diseased cases (the ratings of the non-diseased cases are the FP ratings on diseased cases). Fit RSM to this dataset. Proceed as before. Key assumption being violated: the FP ratings on diseased cases are independent of the TP ratings on same cases. However, without this assumption one cannot estimate RSM parameters. Need `RJafroc` function to handle this special case: `FitRsmRoc1`? No! Just need function to create a "ROC" dataset from one that only has diseased cases. e.g., FixK1eq0dataset?



```{r vary-lambda-parameters, echo=TRUE}
muArr <- c(2)
lambdaArr <- c(1, 2, 5, 10)
nuArr <- c(0.9)
lesDistr <- c(0.5, 0.5)
relWeights <- c(0.5, 0.5)
```



