# How to analyze a dataset with zero non-diseased cases  {#analyze-diseased-only-dataset}

---
output:
  rmarkdown::pdf_document:
    fig_caption: yes        
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval.after = "fig.cap"
)

library(kableExtra)
library(gridExtra)
library(abind)
library(dplyr)
```




## TBA How much finished {#analyze-diseased-only-dataset-how-much-finished}
0%


## The problem {#analyze-diseased-only-dataset-methods}

How to analyze $K_1 = 0$ datasets.


## Handling diseased-only datasets {#diseased-only-dataset}

ROC-like plot of TPF vs. FPF1 is possible. Can create a ROC-like dataset with equal number of "non-diseased" and diseased cases (the ratings of the non-diseased cases are the FP ratings on diseased cases). Fit RSM to this dataset. Proceed as before. Key assumption being violated: the FP ratings on diseased cases are independent of the TP ratings on same cases. However, without this assumption one cannot estimate RSM parameters. Need `RJafroc` function to handle this special case: `FitRsmRoc1`? No! Just need function to create a "ROC" dataset from one that only has diseased cases. e.g., FixK1eq0dataset?


Save TONY dataset to `dsTony`. Create copy `dsNoNormals`. Replace all marks with case indices greater than $K_2$ with `NA`. This creates a dataset with no marks on non-diseased cases.

```{r analyze-diseased-only-code, echo=TRUE}
dsTony <- RJafroc::dataset01 # TONY dataset
K2 <- length(dsTony$lesions$perCase)
K1 <- length(dsTony$ratings$NL[1,1,,1]) - K2
dsNoNormals <- dsTony
dsNoNormals$ratings$NL[,,1:K1,] <- NA # remove any marks from non-diseased cases
UtilFigureOfMerit(dsTony,FOM = "wAFROC")
UtilFigureOfMerit(dsNoNormals,FOM = "wAFROC")
UtilFigureOfMerit(dsTony,FOM = "wAFROC1")
UtilFigureOfMerit(dsNoNormals,FOM = "wAFROC1")
ds1 <- array(NA, dim = c(2,5,K2,3))
for (i in 1:2) {
  for (j in 1:5) {
    
  } 
}
```



