# (PART\*) DATASETS {-}


# Datasets {#datasets}


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
```


## Datasets embedded in `RJafroc` {#datasets-datasetsdd}

They are identified in the code by dataset`dd` (where `dd` is an integer in the range `01` to `14`). As an example, `dataset01` can be viewed [here](https://dpc10ster.github.io/RJafroc/reference/dataset01.html). 


### Dataset01 {#datasets-dataset01}

`dataset01` "TONY" FROC dataset [@chakraborty2011estimating]

```{r, echo=FALSE}
str(RJafroc::dataset01$ratings)
```


### Dataset02 {#datasets-dataset02}

`dataset02` "VAN-DYKE" (Van Dyke) ROC dataset [@RN1993]

```{r, echo=FALSE}
str(RJafroc::dataset02$ratings)
```


### Dataset03 {#datasets-dataset03}

`dataset03` "FRANKEN" (Franken) ROC dataset [@RN1995]

```{r, echo=FALSE}
str(RJafroc::dataset03$ratings)
```


### Dataset04 {#datasets-dataset04}

`dataset04` "FEDERICA" (Federica Zanca) FROC dataset [@zanca2009evaluation]

```{r, echo=FALSE}
str(RJafroc::dataset04$ratings)
```


### Dataset05 {#datasets-dataset05}


`dataset05` "THOMPSON" (John Thompson) FROC dataset [@RN2368]

```{r, echo=FALSE}
str(RJafroc::dataset05$ratings)
```


### Dataset06 {#datasets-dataset06}

* `dataset06` "MAGNUS" (Magnus Bath) FROC dataset [@RN1929]

```{r, echo=FALSE}
str(RJafroc::dataset06$ratings)
```


### Dataset07 {#datasets-dataset07}

`dataset07` "LUCY-WARREN" (Lucy Warren) FROC dataset [@RN2507]

```{r, echo=FALSE}
str(RJafroc::dataset07$ratings)
```


### Dataset08 {#datasets-dataset08}

`dataset08` "PENEDO" (Monica Penedo) FROC dataset [@RN1520]

```{r, echo=FALSE}
str(RJafroc::dataset08$ratings)
```


### Dataset09 {#datasets-dataset09}

`dataset09` "NICO-CAD-ROC" (Nico Karssemeijer) ROC dataset [@hupse2013standalone]

```{r, echo=FALSE}
str(RJafroc::dataset09$ratings)
```


### Dataset10 {#datasets-dataset10}

`dataset10` "RUSCHIN" (Mark Ruschin) ROC dataset [@RN1646]

```{r, echo=FALSE}
str(RJafroc::dataset10$ratings)
```


### Dataset11 {#datasets-dataset11}

`dataset11` "DOBBINS-1" (James Dobbins) FROC dataset [@Dobbins2016MultiInstitutional]

```{r, echo=FALSE}
str(RJafroc::dataset11$ratings)
```


### Dataset12 {#datasets-dataset12}

`dataset12`  "DOBBINS-2" (James Dobbins) ROC dataset [@Dobbins2016MultiInstitutional]

```{r, echo=FALSE}
str(RJafroc::dataset12$ratings)
```


### Dataset13 {#datasets-dataset13}

`dataset13` "DOBBINS-3" (James Dobbins) FROC dataset [@Dobbins2016MultiInstitutional]

```{r, echo=FALSE}
str(RJafroc::dataset13$ratings)
```


### Dataset14 {#datasets-dataset14}


`dataset14` "FEDERICA-REAL-ROC" (Federica Zanca) *real* ROC dataset [@RN2318]


```{r, echo=FALSE}
str(RJafroc::dataset14$ratings)
```


## Other datasets

### DatasetCadLroc {#datasets-datasetCadLroc}

`datasetCadLroc` "NICO-CAD-LROC" (Nico Karssemeijer) standalone CAD LROC dataset


```{r, echo=FALSE}
str(RJafroc::datasetCadLroc$ratings)
```


### datasetCadSimuFroc {#datasets-datasetCadSimuFroc}

`datasetCadSimuFroc` "SIM-CAD-FROC" (Nico Karssemeijer) simulated standalone CAD FROC dataset


```{r, echo=FALSE}
str(RJafroc::datasetCadSimuFroc$ratings)
```

